Model

### 2019-09-19 stats/model.Rmd
clean up model file and add model run

## Model utilities

@@

# Inferring semantic knowledge
```{r}
semStr <- '
var data = dataFromR.data

# Model runs

sem_know_model <- readRDS("../saves/sem_know_model_variant3_100k_log_like.rds")

#saveRDS(sem_know_model, "../saves/sem_know_model_wide_sig_ind.rds")

# Summarizing parameters

```
# Plotting model parameters

## Plotting semantic knowledge
```{r}
aoa_ratings_all <- read_xlsx(path = "../data/words_aoa_ratings.xlsx", sheet = 1)

### Comparing parameters between model runs

#### sem knowledge

```{r}
# me_model_data <- read_csv("../data/me.csv")%>%
#     filter(trial != "train1",
#            trial != "train2")%>%
#   mutate(age = age_num - min(age_num))
# 
# sem_know_run_summary <- item_params%>%
#   spread(parameter,value)%>%
#   group_by(Chain,item)%>%
#   summarise(overall_int = estimate_mode(int),
#             overall_slope = estimate_mode(slope))
# 
# 
# sem_know_run_iteration <- item_params%>%
#   spread(parameter,value)
# 
# sem_know_chain <- sem_know_run_iteration %>%
#   left_join(sem_know_run_summary)
# 
# chain_comp <- sem_know_chain %>%
#   left_join(expand.grid(
#       age = unique(me_model_data$age),
#       item = unique(me_model_data$familiar)), .) %>%
#   mutate(sem_know_ci = plogis(int + slope * age),
#          sem_know = plogis(overall_int + overall_slope * age)) %>%
#   group_by(Chain,item, age) %>%
#   summarise(sem_know = estimate_mode(sem_know), 
#             sem_know_uci = hdi_upper(sem_know_ci),
#             sem_know_lci = hdi_lower(sem_know_ci))
# 
# ggplot(chain_comp, aes(x = age+2, y= sem_know, fill = factor(Chain)))+
#   geom_line(size = .3, aes(col = factor(Chain)))+
#   geom_ribbon(aes(ymin=sem_know_lci, ymax = sem_know_uci), alpha = .05)+
#   ylab("Semantic knowledge")+
#   xlab("Age")+
#   ylim(0,1)+
#   facet_wrap(~item, nrow = 3)+
#   theme_few()+
#   scale_colour_discrete()+
#   scale_fill_discrete()

#ggsave("../graphs/model_by_chain.pdf", width = 5, height = 4, scale = 1.5)
```

#### speaker optimality

```{r}
# me_model_data <- read_csv("../data/me.csv")%>%
#     filter(trial != "train1",
#            trial != "train2")%>%
#   mutate(age = age_num - min(age_num))
# 
# so_run_iteration <- global_params%>%
#   spread(parameter, value)
# 
# so_run_summary <- so_run_iteration%>%
#   group_by(Chain)%>%
#   summarise(overall_int = estimate_mode(int),
#             overall_slope = estimate_mode(slope))
# 
# so_run <- so_run_iteration %>%
#   left_join(so_run_summary)
# 
# so_run_comp <-so_run%>%
#   group_by(Chain,Iteration)%>%
#   tidyr::expand(int,slope,overall_int,overall_slope,age = unique(me_model_data$age))%>%
#   mutate(so_ci = int + slope * age,
#          so = overall_int + overall_slope * age) %>%
#   group_by(Chain, age) %>%
#   summarise(so = estimate_mode(so), 
#             so_uci = hdi_upper(so_ci),
#             so_lci = hdi_lower(so_ci))
# 
# ggplot(so_run_comp, aes(x = age+2, y= so, fill = factor(Chain)))+
#   geom_line(size = .4, aes(col = factor(Chain)))+
#   geom_ribbon(aes(ymin=so_lci, ymax = so_uci), alpha = .05)+
#   ylab("speaker optimality")+
#   xlab("Age")+
#   #facet_grid(~run)+
#   theme_few()+
#   scale_colour_discrete()+
#   scale_fill_discrete()
# 
# ggsave("../graphs/so_by_chain.pdf", width = 4, height = 3, scale = 1.5)
```

#### Correlations between parameters for semantic knowledge

```{r}
# sem_know_model_1<- webppl(
#   program_code = paste(rsaUtils, rsaModel, semStr , sep='\n'),
#   data = list(data = me_model_data),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 10,
#   cores = 10,
#   inference_opts = list(method = "incrementalMH",
#                         samples = 4000,
#                         burn = 2000,
#                         verbose = T,
#                         lag = 1,
#                         verboseLag = 1000)
# )
# 
# sem_know_model_2<- webppl(
#   program_code = paste(rsaUtils, rsaModel, semStr , sep='\n'),
#   data = list(data = me_model_data),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 10,
#   cores = 10,
#   inference_opts = list(method = "incrementalMH",
#                         samples = 4000,
#                         burn = 2000,
#                         verbose = T,
#                         lag = 1,
#                         verboseLag = 1000)
# )%>%
#   mutate(Chain = Chain +10)
# 
# sem_know_model_3<- webppl(
#   program_code = paste(rsaUtils, rsaModel, semStr , sep='\n'),
#   data = list(data = me_model_data),
#   data_var = "dataFromR",
#   model_var = "model",
#   chains = 10,
#   cores = 10,
#   inference_opts = list(method = "incrementalMH",
#                         samples = 4000,
#                         burn = 2000,
#                         verbose = T,
#                         lag = 1,
#                         verboseLag = 1000)
# )%>%
#   mutate(Chain = Chain +20)
# 
# 
# sem_know_cor_params <- bind_rows(
#   sem_know_model_1,
#   sem_know_model_2,
#   sem_know_model_3
# )%>%
#   saveRDS(.,"../saves/run_correlations.rds")

# sem_know_cor_params  <- readRDS("../saves/run_correlations.rds")%>%
#   separate(Parameter, into = c("parameter", "type","item"), sep = ",")%>%
#   filter(parameter == "parameter", type == "items")%>%
#   separate(value, into = c("int", "slope"),sep = ", ")%>%
#   mutate(int = as.numeric(str_remove_all(int, "[(c]")),
#          slope = as.numeric(str_remove(slope, "[)]")))%>%
#   gather(parameter, value, int, slope)
# 
# 
# sem_know_cor_params <- sem_know_cor_params%>%
#   filter(Chain != "15",
#          Chain != "1",
#          Chain != "30",
#          Chain != "6",
#          Chain != "5",
#          Chain != "7",
#          Chain != "3")

# Inverse predictions

```{r predictions pragmatic model ex 3 adults}
inv_model_pred<- webppl(
  program_code = paste(rsaUtils, rsaModel, invPred , sep='\n'),
  data =list(data = inv_model_data),
  data_var = "dataFromR"
)
# 
# #saveRDS(inv_model_pred, "../saves/inv_model_pred.rds")
# #inv_model_pred <- readRDS("../saves/inv_model_pred.rds")
# 
inv_predictions <- inv_model_pred %>%
  separate(`0`, into = c("item", "age"), sep="/")%>%
  mutate(age = as.numeric(age))%>%
  rename( me_effect = `1`)%>%
  left_join(aoa_ratings)%>%
  ungroup()%>%
  mutate(item = fct_reorder(factor(item), mean_aoa),
         age = age + min(me_data$age_num))
```

## For comparison: GLMM for ME effect and data
```{r}
me_data <- read_csv("../data/me.csv")%>%
    filter(trial != "train1",
           trial != "train2")%>%
  mutate(z_age = age_num - mean(age_num))

## Model predictions for combination

comb_model_predictions <- comb_model_pred %>%
  separate(`0`, into = c("alignment","item", "age"), sep="/")%>%
  mutate(age = as.numeric(age))%>%
  rename( me_effect = `1`)%>%
  left_join(aoa_ratings)%>%
  ungroup()%>%
  mutate(item = fct_reorder(factor(item), mean_aoa),
         age = age + min(me_data$age_num),
         source = "Model")

```{r}
ggsave("../graphs/combination_model_data_by_item.pdf", width = 8, height = 3, scale = 1.5)
```
# Correlate model predictions and data 
```{r}
binnned_data <- data_comb%>%
  group_by(subage, alignment, item)%>%
  summarize(k = sum(correct), n = n())%>%
  ungroup() %>%
  mutate(a = 1 + k,
         b = 1 + n - k,
         data_mean = (a-1)/(a+b-2),
         data_ci_lower  = qbeta(.025, a, b),
         data_ci_upper = qbeta(.975, a, b),
         subage = factor(subage))%>%
  select(-a,-b,-n,-k)

# Nesting within participant

@@

## Pragmatics model

@@

@@

sem_know_model<- webppl(
  program_code = paste(rsaUtils, rsaModel, sem_subj_str , sep='\n'),
  data = list(data = me_model_data, comb_data = data_comb_log),
  data_var = "dataFromR",
  model_var = "model",
  chains = 3,
  cores = 3,
  inference_opts = list(method = "incrementalMH",
                        samples = 100000,
                        burn = 50000,
                        verbose = T,
                        lag = 1,
                        verboseLag = 5000)
)

@@

@@

log_like<- sem_know_model %>%
  separate(Parameter, into = c("parameter", "model","age","alignment"), sep = ",")%>%
  filter(parameter == "log_like")%>%
  separate(value, into = c("na", "log_like"),sep = ",")%>%
  select(-na, -age, -alignment)%>%
  mutate(log_like =as.numeric(str_remove(log_like,"[)]")))%>%
  group_by(model, Chain)%>%
  summarize(logP = logSumExp(log_like))

@@

#### plotting predictions

```{r}
comb_model_pred_agg <- model_pred%>%
  group_by(Chain, alignment,item,age)%>%
  summarise(mean = estimate_mode(me_effect),
            uci = hdi_upper(me_effect),
            lci = hdi_lower(me_effect))%>%
  left_join(aoa_ratings) %>%
  ungroup()%>%
  mutate(item = fct_reorder(factor(item), mean_aoa),
         age = age + min(me_data$age_num),
         Chain = factor(Chain))

@@

>>>>

# Model utilities

@@

# Model (nesting within participant)

## Prepare data

@@

## Model code

@@

## Model run

@@

sem_know_model<- webppl(
  program_code = paste(rsaUtils, rsaModel, sem_subj_str , sep='\n'),
  data = list(data = me_model_data, comb_data = data_comb_log),
  data_var = "dataFromR",
  model_var = "model",
  chains = 3,
  cores = 3,
  inference_opts = list(method = "incrementalMH",
                        samples = 100000,
                        burn = 50000,
                        verbose = T,
                        lag = 1,
                        verboseLag = 1000)
)

@@

sem_know_model <- readRDS("../saves/sem_know_model_variant3_100k_model_comparison.rds")

## Summarizing parameters

@@

```
## Plotting model parameters

### Semantic knowledge

### Global parameters

### Global sigmas

@@

## Plotting semantic knowledge

```{r}
aoa_ratings_all <- read_xlsx(path = "../data/words_aoa_ratings.xlsx", sheet = 1)

# MAP based inverse predictions for ME data

```{r predictions pragmatic model ex 3 adults}
inv_model_pred<- webppl(
  program_code = paste(rsaUtils, rsaModel, invPred , sep='\n'),
  data =list(data = inv_model_data),
  data_var = "dataFromR"
)
# 
# #saveRDS(inv_model_pred, "../saves/inv_model_pred.rds")
# #inv_model_pred <- readRDS("../saves/inv_model_pred.rds")
# 
inv_predictions <- inv_model_pred %>%
  separate(`0`, into = c("item", "age"), sep="/")%>%
  mutate(age = as.numeric(age))%>%
  dplyr::rename( me_effect = `1`)%>%
  left_join(aoa_ratings)%>%
  ungroup()%>%
  mutate(item = fct_reorder(factor(item), mean_aoa),
         age = age + min(me_data$age_num))
```

## For comparison: GLMM for ME effect and data

```{r}
me_data <- read_csv("../data/me.csv")%>%
    filter(trial != "train1",
           trial != "train2")%>%
  mutate(z_age = age_num - mean(age_num))

# MAP based Model predictions for combination

comb_model_predictions <- comb_model_pred %>%
  separate(`0`, into = c("alignment","item", "age"), sep="/")%>%
  mutate(age = as.numeric(age))%>%
  dplyr::rename( me_effect = `1`)%>%
  left_join(aoa_ratings)%>%
  ungroup()%>%
  mutate(item = fct_reorder(factor(item), mean_aoa),
         age = age + min(me_data$age_num),
         source = "Model")

```{r}
ggsave("../graphs/combination_model_data_by_item.pdf", width = 8, height = 3, scale = 1.5)
```

## Correlate model predictions and data

```{r}
binnned_data <- data_comb%>%
  group_by(subage, alignment, item)%>%
  summarize(k = sum(correct), n = n())%>%
  ungroup() %>%
  mutate(a = 1 + k,
         b = 1 + n - k,
         data_mean = (a-1)/(a+b-2),
         data_ci_lower  = qbeta(.025, a, b),
         data_ci_upper = qbeta(.975, a, b),
         subage = factor(subage))%>%
  select(-a,-b,-n,-k)

## Model compatison

```{r}
log_like<- sem_know_model %>%
  separate(Parameter, into = c("parameter", "model","age","alignment"), sep = ",")%>%
  filter(parameter == "log_like")%>%
  separate(value, into = c("na", "log_like"),sep = ",")%>%
  select(-na, -age, -alignment)%>%
  mutate(log_like =as.numeric(str_remove(log_like,"[)]")))%>%
  group_by(model,Chain)%>%
  summarize(logP = logSumExp(log_like))
```

## Plotting model predictions

```{r}
model_pred<- sem_know_model %>%
  separate(Parameter, into = c("parameter", "item","age","alignment"), sep = ",")%>%
  filter(parameter == "prediction")%>%
  separate(value, into = c("na", "me_effect"),sep = ",")%>%
  select(-na)%>%
  mutate(age = as.numeric(age),
         me_effect =as.numeric(str_remove(me_effect,"[)]")))

comb_model_pred_agg <- model_pred%>%
  group_by(Chain, alignment,item,age)%>%
  summarise(mean = estimate_mode(me_effect),
            uci = hdi_upper(me_effect),
            lci = hdi_lower(me_effect))%>%
  left_join(aoa_ratings) %>%
  ungroup()%>%
  mutate(item = fct_reorder(factor(item), mean_aoa),
         age = age + min(me_data$age_num),
         Chain = factor(Chain))

@@

# OLD: Comparing parameters between model runs

## sem knowledge

sem_know_run_summary <- item_params%>%
  spread(parameter,value)%>%
  group_by(Chain,item)%>%
  summarise(overall_int = estimate_mode(int),
            overall_slope = estimate_mode(slope))

sem_know_run_iteration <- item_params%>%
  spread(parameter,value)

sem_know_chain <- sem_know_run_iteration %>%
  left_join(sem_know_run_summary)

chain_comp <- sem_know_chain %>%
  left_join(expand.grid(
      age = unique(me_model_data$age),
      item = unique(me_model_data$familiar)), .) %>%
  mutate(sem_know_ci = plogis(int + slope * age),
         sem_know = plogis(overall_int + overall_slope * age)) %>%
  group_by(Chain,item, age) %>%
  summarise(sem_know = estimate_mode(sem_know),
            sem_know_uci = hdi_upper(sem_know_ci),
            sem_know_lci = hdi_lower(sem_know_ci))

ggplot(chain_comp, aes(x = age+2, y= sem_know, fill = factor(Chain)))+
  geom_line(size = .3, aes(col = factor(Chain)))+
  geom_ribbon(aes(ymin=sem_know_lci, ymax = sem_know_uci), alpha = .05)+
  ylab("Semantic knowledge")+
  xlab("Age")+
  ylim(0,1)+
  facet_wrap(~item, nrow = 3)+
  theme_few()+
  scale_colour_discrete()+
  scale_fill_discrete()

## speaker optimality

so_run_iteration <- global_params%>%
  spread(parameter, value)

so_run_summary <- so_run_iteration%>%
  group_by(Chain)%>%
  summarise(overall_int = estimate_mode(int),
            overall_slope = estimate_mode(slope))

so_run <- so_run_iteration %>%
  left_join(so_run_summary)

so_run_comp <-so_run%>%
  group_by(Chain,Iteration)%>%
  tidyr::expand(int,slope,overall_int,overall_slope,age = unique(me_model_data$age))%>%
  mutate(so_ci = int + slope * age,
         so = overall_int + overall_slope * age) %>%
  group_by(Chain, age) %>%
  summarise(so = estimate_mode(so),
            so_uci = hdi_upper(so_ci),
            so_lci = hdi_lower(so_ci))

ggplot(so_run_comp, aes(x = age+2, y= so, fill = factor(Chain)))+
  geom_line(size = .4, aes(col = factor(Chain)))+
  geom_ribbon(aes(ymin=so_lci, ymax = so_uci), alpha = .05)+
  ylab("speaker optimality")+
  xlab("Age")+
  #facet_grid(~run)+
  theme_few()+
  scale_colour_discrete()+
  scale_fill_discrete()

## Correlations between parameters for semantic knowledge

# Model (No nesting within participants)
```{r}
semStr <- '
var data = dataFromR.data
